Neural Power Unit \\-- A Bayesian Approach To Neural Arithmetic

### 2020-05-16 papers/main.tex
update plots packages; figure fuzzing

% figures
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds,positioning,intersections}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{polar}
\usepgfplotslibrary{smithchart}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{dateplot}
\graphicspath{{./../../plots/}}
\newcommand{\real}{\text{real}}

@@

\begin{abstract}
  Common Neural Networks can approximate simple arithmetic operations, but fail
  to generalize well beyond the training manifold. A new class of units called
  \emph{Neural Arithmetic Units} aim to overcome this difficulty, but are limited
  either to operate on positive numbers (NALU by
  \citet{trask_neural_2018}), or can only represent simple addition and
  multiplication (NAU \& NMU by \citet{madsen_neural_2020}).
  We present the first Neural Arithmetic Unit that operates on the full domain
  of real numbers and is capable of learning arbitrary power functions.
\end{abstract}

@@

Numbers and simple algebra are essential not only to human intelligence but also
to the survival of many other
species~\citep{dehaene_number_2011,gallistel_finding_2018}.  This can be taken
as a hint that arithmetic is an important ingredient to a successful,
intelligent agent.  State of the art neural networks are capable of learning
simple arithmetic, but they fail to extrapolate beyond the ranges seen during
training~\citep{suzgun_evaluating_2018,lake_generalization_2018}.
The inability to generalize to unseen inputs is a fundamental problem that
hints at a lack of \emph{understanding} of the given task. The model merely
memorizes the seen inputs and fails to abstract the true learning task.
The failure of numerical extrapolation on simple arithmetic tasks has been shown
by~\cite{trask_neural_2018},
who also introduce a new class of \emph{Neural Arithmetic Units}
that show exceptional extrapolation performance on most arithmetic tasks.
...now explain drawbacks of NALU and NMU...
...we introduce blah...

  This
can be demonstrated by training a traditional, dense NN to approximate three
functions
\begin{align}
  \label{eq:approx_tasks}
  f(x) = e^x, && g(x) = \log(x), && h(x) = \sin(x)
\end{align}

\begin{figure}
  \centering
  \input{../plots/exp_log_sin/gatednpux_exp.tikz}
  \caption{}%
  \label{fig:}
\end{figure}

@@

@@

\begin{align}
  \label{eq:nalu_add}
  \text{Addition: }       & \bm a = \bm W \bm x
                          & \bm W& = \tanh(\hat{\bm W}) \odot \sigma(\hat{\bm M}) \\
  \label{eq:nalu_mult}
  \text{Multiplication: } & \bm m = \exp \bm W(\log(|\bm x|+\epsilon)) & &\\
  \text{Output: }         & \bm y = \bm a \odot \bm g + \bm m \odot (1-\bm g) 
                          & \bm g& = \sigma(\bm G\bm x)
\end{align}

A multiplication layer that can handle negative inputs was introduced by \citet{madsen_neural_2020}.
The \emph{Neural Multiplication Unit} (NMU) is defined by Eq.~\ref{eq:nmu} and is typically used in
conjunction with the so-called \emph{Neural Addition Unit} (NAU) in Eq.~\ref{eq:nau}.
\begin{align}
  \label{eq:nmu}
  \text{NMU: } &y_j = \prod_i M_{ij} z_{i} + 1 - M_{ij}  &M_{ij}=\min(\max(M_{ij}, 0), 1)\\
  \label{eq:nau}
  \text{NAU: } &\bm y = \bm A \bm x &A_{ij}=\min(\max(A_{ij}, 0), 1)
\end{align}
In both NMU and NAU the weights are clipped to $[0,1]$, and typically regularized
with $\mathcal{R}$:
\begin{align}
  \label{eq:rsparse}
  \mathcal{R} = \sum_{ij} \min(W_{ij}, 1-W_{ij})
\end{align}

@@

\todo{put this in acknowledgements
all the results in this paper were create with the help of the following Julia
packages: \cite{rackauckas_differentialequationsjl_2017} + (Flux.jl)}
\todo{add sqrt and positive range to table~\ref{tab:compare_npu_npu_nalu}?}
\begin{table}
  \centering
  \caption{Comparison of different $2\times2$ layers on the task
  $(x,y)\rightarrow(xy,\frac{x}{y})$.  $(x,y) \in \mathcal U^2(-2,2)$.}
  \label{tab:compare_npu_npu_nalu}
  \input{table_x*y_xdivy.tex}
\end{table}

@@

\section{Experimetns}%
\label{sec:experimetns}

>>>>

% figures
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds,positioning,intersections}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{polar}
\usepgfplotslibrary{smithchart}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{dateplot}
\graphicspath{{./../plots/}}
\newcommand{\real}{\text{real}}

@@

\begin{abstract}
  Common Neural Networks can approximate simple arithmetic operations, but fail
  to generalize beyond the range of numbers that were seen during training. A
  new class of units called \emph{Neural Arithmetic Units} aim to overcome this
  difficulty, but are limited either to operate on positive numbers (NALU by
  \citet{trask_neural_2018}), or can only represent simple addition and
  multiplication (NAU \& NMU by \citet{madsen_neural_2020}).  We introduce a
  Neural Arithmetic Unit that operates on the full domain of real numbers and
  is capable of learning arbitrary power functions called \emph{Neural Power
  Unit} (NPU). The NPU enables not only the learning simple arithmetic
  operations, but is also capable of learning e.g. Taylor expansions with very
  few parameters, which makes the NPU highly explainable.
  \footnote{Implementation of neural arithmetic units:
  \url{https://github.com/nmheim/NeuralArithmetic.jl}}.
\end{abstract}

@@

Numbers and simple algebra are essential not only to human intelligence but
also to the survival of many other
species~\citep{dehaene_number_2011,gallistel_finding_2018}.  This can be taken
as a hint that arithmetic is an important ingredient to a successful,
intelligent agent.  State of the art neural networks are capable of learning
simple arithmetic, but they fail to extrapolate beyond the ranges seen during
training~\citep{suzgun_evaluating_2018,lake_generalization_2018}.  The
inability to generalize to unseen inputs is a fundamental problem that hints at
a lack of \emph{understanding} of the given task. The model merely memorizes
the seen inputs and fails to abstract the true learning task.  The failure of
numerical extrapolation on simple arithmetic tasks has been shown
by~\cite{trask_neural_2018}, who also introduce a new class of \emph{Neural
Arithmetic Units} that show good extrapolation performance on some arithmetic
tasks.

The \emph{Neural Arithmetic Logic Unit} (NALU) can learn addition,
subtraction, multiplication, and division with two severe limitations: the
inputs that the NALU is trained with have to be positive, and should be
significantly larger than zero, otherwise the layer will not converge to
a correct solution.

\cite{madsen_neural_2020} introduce two new arithmetic layers, the \emph{Neural
Addition Unit} (NAU) and the \emph{Neural Multiplication Unit} (NMU).
Stacked, they can learn addition, subtraction, and multiplication and they converge
for any input range. However, the NMU cannot learn division.

\subsection*{Our Contribution}%
\label{sub:our_contribution}

We introduce a new arithmetic layer called the \emph{Neural Power Unit} (NPU)
which is capable of learning addition, subtraction, and arbitrary power
functions, which includes multiplication and division.
Our approach is inspired by the NALU, but overcomes its limitations to
non-small positive numbers by leveraging the complex plane.

\begin{figure}
  \centering
  %\includegraphics[width=0.8\linewidth]{compare-npu-nmu-nalu-xy-xdivy.tikz}
  \input{../plots/compare-npu-nmu-nalu-xy-xdivy.tex}
  \caption{Comparison of different models learning the function
  $(x,y)\rightarrow(xy,\frac{x}{y})$ with $(x,y)\in\mathcal U^2(0.1,2)$.  Each
  plot shows the mean-squared error on a logarithmic color scale ($\hat{\bm
  t}=\text{model}(\bm x)$).  The Dense network learns both multiplication and
  division, but fails to extrapolate. NALU learns both tasks perfectly, but
  completely fails if one of the inputs is negative.  The NMU successfully
  learns multiplication, but not division. Only the NPU learns both tasks and
  can extrapolate with an error that is less than $0.01$ in most places.}%
  \label{fig:compare-npu-nmu-nalu-xy-xdivy}
\end{figure}

@@

\subsection{Neural Arithmetic Logic Unit}%
\label{sub:neural_arithmetic_logic_unit}

@@

\begin{align}
  \label{eq:nalu_add}
  \text{Addition: }       & \bm a = \bm W \bm x
                          & \bm W& = \tanh(\hat{\bm W}) \odot \sigma(\hat{\bm M}) \\
  \label{eq:nalu_mult}
  \text{Multiplication: } & \bm m = \exp \bm W(\log(|\bm x|+\epsilon)) & &\\
  \text{Output: }         & \bm y = \bm a \odot \bm g + \bm m \odot (1-\bm g) 
                          & \bm g& = \sigma(\bm G\bm x)
\end{align}
Additionally the NALU has severe convergence problems for small inputs due to
a non-smooth loss surface close to zero, as shown by \cite{madsen_neural_2020}.

\subsection{Neural Multiplication Unit}%
\label{sub:neural_multiplication_unit}

A multiplication layer that can handle small and negative inputs was introduced
by \citet{madsen_neural_2020}.  The \emph{Neural Multiplication Unit} (NMU) is
defined by Eq.~\ref{eq:nmu} and is typically used in conjunction with the
so-called \emph{Neural Addition Unit} (NAU) in Eq.~\ref{eq:nau}.
\begin{align}
  \label{eq:nmu}
  \text{NMU: } &y_j = \prod_i M_{ij} z_{i} + 1 - M_{ij}  &M_{ij}=\min(\max(M_{ij}, 0), 1)\\
  \label{eq:nau}
  \text{NAU: } &\bm y = \bm A \bm x &A_{ij}=\min(\max(A_{ij}, 0), 1)
\end{align}
In both NMU and NAU the weights are clipped to $[0,1]$, and typically regularized
with $\mathcal{R}$:
\begin{align}
  \label{eq:rsparse}
  \mathcal{R} = \sum_{ij} \min(W_{ij}, 1-W_{ij})
\end{align}

@@

\todo{put this in acknowledgements
all the results in this paper were create with the help of the following Julia
packages: \cite{rackauckas_differentialequationsjl_2017} + (Flux.jl)}
\todo{add sqrt and positive range to table~\ref{tab:compare_npu_npu_nalu}?}
\begin{table}
  \centering
  \caption{Comparison of different $2\times2$ layers on the task
  $(x,y)\rightarrow(xy,\frac{x}{y})$.  $(x,y) \in \mathcal U^2(-2,2)$.}
  \label{tab:compare_npu_npu_nalu}
  \input{table-x*y-xdivy-pos-neg.tex}
\end{table}

@@

\section{Experimetns}%
\label{sec:experimetns}
This can be demonstrated by training a traditional, dense NN to approximate
three functions
\begin{align}
  \label{eq:approx_tasks}
  f(x) = e^x, && g(x) = \log(x), && h(x) = \sin(x)
\end{align}
